### 2025-04-04 about/vita.tex
uncertainty acceptance

\begingroup\frenchspacing
\papercount=6
\pubbegin P
\papitem (with Jonah Saks)
On the homology of several number-theoretic set families.
{\sl Enumerative Combinatorics and Applications}\/ {\bf 4},2 (2024), \#S2R12, 11~pp.
[arXiv:2206.12535]
\papitem (with Jad Hamdan and Jonah Saks)
The lattice of arithmetic progressions.
{\sl Australasian Journal of Combinatorics}\/ {\bf 84},3 (2022), 357--374.
[arXiv:2106.05949]
\papitem (with Luc Devroye and Rosie Y. Zhao)
On the peel number and the leaf-height of a Galton--Watson tree.
{\sl Combinatorics, Probability and Computing}\/ {\bf 32},1 (2023), 68--90.
[arXiv:2106.14389]
\papitem (with Anna M. Brandenberger, Luc Devroye, and Rosie Y. Zhao)
Leaf multiplicity in a Bienaym\'e--Galton--Watson tree. {\sl Discrete Mathematics
and Theoretical Computer Science}\/ {\bf 24},1 (2022), \#7, 16~pp.
[arXiv:2105.12046]
\papitem (with Anna M. Brandenberger and Luc Devroye)
Root estimation in Galton--Watson trees. {\sl Random Structures and Algorithms}\/ {\bf 61},3 (2022), 520--542.
[arXiv:2007.05681]
\papitem (with Rosie Y. Zhao)
Arithmetic subsequences in a random ordering of an additive set.
{\sl Integers: Electronic Journal of Combinatorial Number Theory}\/ {\bf 21} (2021), \#A89, 19~pp.
[arXiv:2012.12339]

@@

\fictioncount=4
\pubbegin F
\starficitem ``Lanterns,'' {\sl The Malahat Review} {\bf 229} (2025), 7--13.
Winner of the 2024 Constance Rooke Creative Nonfiction Prize.
\ficitem ``Senang Diri,'' {\sl Existere} {\bf 42},2 (2023), 17--25.
\ficitem ``The Vigil,'' {\sl Ricepaper Magazine} (2023), online. Reprinted in
{\sl Best Canadian Stories 2025} (Windsor, Ont.: Biblioasis, 2024), 244--253.
\ficitem ``Mountain Pass,'' {\sl Prairie Journal of Canadian Literature} {\bf 79} (2023), 48--55.
\endgroup%end the frenchspacing

@@

\starleftright{\bf Course lecturer}{\bf August 2024 -- present}
\leftright{\sl McGill University}{\sl Montr\'eal, QC, Canada}
\smallskip
Instructor for a section of MATH 240 Discrete Structures. Duties include preparing and delivering lectures,
designing assignment and exam questions, and holding office hours.
\medbreak

@@

>>>>

\begingroup\frenchspacing
\papercount=7
\pubbegin P
\papitem An uncertainty principle for M\"obius inversion on posets.
To appear in {\sl Contributions to Discrete Mathematics}.
[arXiv:2302.02466]
\papitem (with Jonah Saks)
On the homology of several number-theoretic set families.
{\sl Enumerative Combinatorics and Applications}\/ {\bf 4},2 (2024), \#S2R12, 11~pp.
[arXiv:2206.12535]
\papitem (with Jad Hamdan and Jonah Saks)
The lattice of arithmetic progressions.
{\sl Australasian Journal of Combinatorics}\/ {\bf 84},3 (2022), 357--374.
[arXiv:2106.05949]
\papitem (with Luc Devroye and Rosie Y. Zhao)
On the peel number and the leaf-height of a Galton--Watson tree.
{\sl Combinatorics, Probability and Computing}\/ {\bf 32},1 (2023), 68--90.
[arXiv:2106.14389]
\papitem (with Anna M. Brandenberger, Luc Devroye, and Rosie Y. Zhao)
Leaf multiplicity in a Bienaym\'e--Galton--Watson tree. {\sl Discrete Mathematics
and Theoretical Computer Science}\/ {\bf 24},1 (2022), \#7, 16~pp.
[arXiv:2105.12046]
\papitem (with Anna M. Brandenberger and Luc Devroye)
Root estimation in Galton--Watson trees. {\sl Random Structures and Algorithms}\/ {\bf 61},3 (2022), 520--542.
[arXiv:2007.05681]
\papitem (with Rosie Y. Zhao)
Arithmetic subsequences in a random ordering of an additive set.
{\sl Integers: Electronic Journal of Combinatorial Number Theory}\/ {\bf 21} (2021), \#A89, 19~pp.
[arXiv:2012.12339]

@@

\fictioncount=4
\pubbegin F
\starficitem ``Lanterns,'' {\sl The Malahat Review} {\bf 229} (2025), 7--13.
Winner of the 2024 Constance Rooke Creative Nonfiction Prize.
\ficitem ``Senang Diri,'' {\sl Existere} {\bf 42},2 (2023), 17--25.
\ficitem ``The Vigil,'' {\sl Ricepaper Magazine} (2023), online. Reprinted in
{\sl Best Canadian Stories 2025} (Windsor, Ont.: Biblioasis, 2024), 244--253.
Reprinted in {\sl Infusion} (Toronto, Ont.: Dark Helix Press, 2025), 37--47.
\ficitem ``Mountain Pass,'' {\sl Prairie Journal of Canadian Literature} {\bf 79} (2023), 48--55.
\endgroup%end the frenchspacing

@@

@@

\starleftright{\bf Course lecturer}{\bf August -- December 2024}
\leftright{\sl McGill University}{\sl Montr\'eal, QC, Canada}
\smallskip
Instructor for a section of MATH 240 Discrete Structures. Duties included preparing and delivering lectures,
designing assignment and exam questions, and holding office hours.
\medbreak

### 2025-04-04 misc/expo/entropy.tex
uncertainty acceptance

On the other hand, say $|A| = m$ and $|B| = n$ with $m<n$. If $m=1$, then by the previous proposition
we have $\Eta\{X\} = 0$, and by normalisation and invariance, $\Eta\{Y\} = 1$.
When $m\ge 2$, pick $k$ such that $m^k \le n^{k-1}$, so that $|A^k| \le |B^{k-1}|$.
Then by Corollary~{\corxn} and the inequality we showed in the first paragraph of this proof, we have
$$n\Eta\{X\} = \Eta\{X^n\} \le \Eta\{X^{n-1}\} = (n-1)\Eta\{Y\},$$
whence $\Eta\{X\} < \Eta\{Y\}$.\slug

@@

\proof Let $A$ be the support of $X$ and $B$ be the support of $Y$.
First we consider the case that $X$ is uniform on $A$ (so $A$ is finite). Then by the definition
of conditional entropy,
$$\Eta\{X\given Y\} = \sum_{b\in B} \pr\{Y = b\} \Eta\{X \given Y= b\}.$$
But for each $b$, the random variable $(X\given Y = b)$ takes values in $A$, so its entropy is
bounded above by $\Eta\{X\}$ by the maximisation axiom. Hence $\Eta\{X\given Y\} \le \Eta\{X\}$.

>>>>

On the other hand, say $|A| = m$ and $|B| = n$ with $m<n$. We want to conclude the
strict inequality $\Eta\{X\} < \Eta\{Y\}$.
Pick $k$ such that $m^k \le n^{k-1}$, so that $A^k$ can be embedded (as a set) into $B^{k-1}$.
By invariance we can view $X^k$ as a random variable that is uniform on a subset of $B^{k-1}$.
so by the inequality we showed in the first paragraph of this proof, as well as Corollary~{\corxn},
we have
$$n\Eta\{X\} = \Eta\{X^n\} \le \Eta\{Y^{n-1}\} = (n-1)\Eta\{Y\},$$
whence $\Eta\{X\} < \Eta\{Y\}$.\slug

The argument in the last paragraph of the previous proof is an instance of the ``tensor power trick,''
in which one amplifies a known inequality by applying it to suitably high powers of the objects
involved.

@@

\proof Let $A$ be the support of $X$ and $B$ be the support of $Y$.
First we consider the case that $X$ is uniform on $A$ (so $A$ is finite). Then by the definition
of conditional entropy,
$$\Eta\{X\given Y\} = \sum_{b\in B} \pr\{Y = b\} \Eta\{X \given Y= b\}.$$
But for each $b$, the random variable $(X\given Y = b)$ takes values in $A$, so its entropy is
bounded above by $\Eta\{X\}$ by the maximisation axiom. Hence $\Eta\{X\given Y\} \le \Eta\{X\}$,
which is equivalent to $\Eta\{Y\given X\} \le \Eta\{Y\}$.
